High-level cognition is supported by at least second order dynamic correlations in neural activity patterns

### 2019-09-01 paper/main.tex
changed figure 2 caption to match methods section text

\subsection*{Kernel-based approach for computing dynamic correlations}
Given a matrix of observations, we can compute the (static)
Pearson's correlation between any pair of columns, $\mathbf{X}(\cdot, i)$ and
$\mathbf{X}(\cdot, j)$ using~\citep{Pear01}:
\begin{align}
  \mathrm{corr}(\mathbf{X}(\cdot, i), \mathbf{X}(\cdot, j)) &=
                                                              \frac{\sum_{t=1}^T
                                                              \left(\mathbf{X}(t,
                                                              i)
                                                              -
                                                              \bar{\mathbf{X}(\cdot,
                                                              i)}\right)
                                                              \left(\mathbf{X}(t,
                                                              j)
                                                              -
                                                              \bar{\mathbf{X}(\cdot, j)}\right)}{\sqrt{\sum_{t=1}^T
                                                              \sigma^2_{\mathbf{X}(\cdot, i)} 
                                                              \sigma^2_{\mathbf{X}(\cdot, j)}}},~\mathrm{where}\\\label{eqn:corr}
  \bar{\mathbf{X}(\cdot, k)} &= \sum_{t=1}^T
                       \mathbf{X}(t, k),~\mathrm{and}\\
  \sigma^2_{\mathbf{X}(\cdot, k)} &= \sum_{t=1}^T \left( \mathbf{X}(t, k) -
                            \bar{\mathbf{X}(\cdot, k)} \right)^2 
\end{align}
We can generalize this formula to compute time-varying correlations by
incorporating a \textit{kernel function} that takes a time $t$ as
input, and returns how much the observed data at each timepoint
$\tau \in \left[ -\infty, \infty \right]$ contributes to the estimated instantaneous
correlation at time $t$ (Fig.~\ref{fig:kernels}).
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/kernels}
  \caption{\textbf{Examples of kernel functions.} Each panel
    displays per-timepoint weights at $t = 50$, evaluated for 100
    timepoints ($\tau \in \left[1, ..., 100\right]$).  \textbf{a. Uniform
      kernel.} The weights are timepoint-invariant; observations at
    all timepoints are weighted equally, and do not change as a
    function of $\tau$.  This is a special case of weight function that
    reduces dynamic correlations to static correlations.
    \textbf{b. Dirac $\delta$ kernel.} Only the observation at
    timepoint $t$ is given a non-zero weight (of 1).
    \textbf{c. Gaussian kernels.} Each kernel's weights fall off
    in time according to a Gaussian probability density function
    centered on time $t$.  Weights derived using several different
    example variance parameters ($\sigma^2$ are displayed.
    \textbf{d. Laplace kernels.}  Each kernel's weights fall off
    in time according to a Laplace probability density function
    centered on time $t$.  Weights derived using several different
    example scale parameters ($b$) are displayed.  \textbf{e. Mexican
      hat (Ricker wavelet) kernels.}  Each kernel's weights fall
    off in time according to a Ricker wavelet centered on time $t$.  This
    function highlights the \textit{contrasts} between local versus
    surrounding activity patterns in estimating dynamic
    correlations. Weights derived using several different example
    width parameters ($\sigma$) are displayed.}
  \label{fig:kernels}
\end{figure}

>>>>

\subsection*{Kernel-based approach for computing dynamic correlations}
Given a matrix of observations, we can compute the (static)
Pearson's correlation between any pair of columns, $\mathbf{X}(\cdot, i)$ and
$\mathbf{X}(\cdot, j)$ using~\citep{Pear01}:
\begin{align}
  \mathrm{corr}(\mathbf{X}(\cdot, i), \mathbf{X}(\cdot, j)) &=
                                                              \frac{\sum_{t=1}^T
                                                              \left(\mathbf{X}(t,
                                                              i)
                                                              -
                                                              \bar{\mathbf{X}(\cdot,
                                                              i)}\right)
                                                              \left(\mathbf{X}(t,
                                                              j)
                                                              -
                                                              \bar{\mathbf{X}(\cdot, j)}\right)}{\sqrt{\sum_{t=1}^T
                                                              \sigma^2_{\mathbf{X}(\cdot, i)} 
                                                              \sigma^2_{\mathbf{X}(\cdot, j)}}},~\mathrm{where}\\\label{eqn:corr}
  \bar{\mathbf{X}(\cdot, k)} &= \sum_{t=1}^T
                       \mathbf{X}(t, k),~\mathrm{and}\\
  \sigma^2_{\mathbf{X}(\cdot, k)} &= \sum_{t=1}^T \left( \mathbf{X}(t, k) -
                            \bar{\mathbf{X}(\cdot, k)} \right)^2 
\end{align}
We can generalize this formula to compute time-varying correlations by
incorporating a \textit{kernel function} that takes a time $t$ as
input, and returns how much the observed data at each timepoint
$\tau \in \left[ -\infty, \infty \right]$ contributes to the estimated instantaneous
correlation at time $t$ (Fig.~\ref{fig:kernels}).
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/kernels}
  \caption{\textbf{Examples of kernel functions.} Each panel
    displays per-timepoint weights at $t = 50$, evaluated for 100
    timepoints ($\tau \in \left[1, ..., 100\right]$).  \textbf{a. Uniform
      kernel.} The weights are timepoint-invariant; observations at
    all timepoints are weighted equally, and do not change as a
    function of $\tau$.  This is a special case of weight function that
    reduces dynamic correlations to static correlations.
    \textbf{b. Dirac $\delta$ kernel.} Only the observation at
    timepoint $t$ is given a non-zero weight (of 1).
    \textbf{c. Gaussian kernels.} Each kernel's weights fall off
    in time according to a Gaussian probability density function
    centered on time $t$.  Weights derived using several different
    example width parameters ($\sigma^2$) are displayed.
    \textbf{d. Laplace kernels.}  Each kernel's weights fall off
    in time according to a Laplace probability density function
    centered on time $t$.  Weights derived using several different
    example width parameters ($b$) are displayed.  \textbf{e. Mexican
      hat (Ricker wavelet) kernels.}  Each kernel's weights fall
    off in time according to a Ricker wavelet centered on time $t$.  This
    function highlights the \textit{contrasts} between local versus
    surrounding activity patterns in estimating dynamic
    correlations. Weights derived using several different example
    width parameters ($\sigma$) are displayed.}
  \label{fig:kernels}
\end{figure}
